---
title: "MethylCallR_address_chain_file_and_array_conversion"
author: "Hyun-Ho, Yang"
date: "`r Sys.Date()`"
output: html_document
---


## Summary

This example explains the method to control duplicated probes, generate the address chain file, and convert EPICv2 data to previous versions.

To provide an example, we used the demo data downloaded from the Illumina official website. The data can be accessed at \url{https://support.illumina.com/array/array_kits/infinium-methylationepic-beadchip-kit/downloads.html}.

## Data preparation

The downloaded raw IDAT files should be unzipped and placed in the same directory. In this example, we assumed that the raw IDAT files were stored in "`/user/MethylCallR_Test/Illumina_demo_EPICV2/`".

When importing demo data through MethylCallR, the required sample data file can be found in the directory `MethylCallR/doc/MethylCallR_address_chain_file_and_array_conversion/MethylCallR_address_chain_file_and_array_conversion_Files`.

We assumed that the sample data file was saved in "`/user/MethylCallR_Test/Illumina_demo_EPICV2/EPICv2_sample_data.csv`".


```{r Chunk1, include=FALSE}
data.Import <- MeCall.ReadIdat(idat.dir = "/user/MethylCallR_Test/Illumina_demo_EPICV2/", pd.file = "/user/MethylCallR_Test/Illumina_demo_EPICV2/EPICv2_sample_data.csv", offset = 100, platform = c("EPICv2"))
```

```{r Chunk1-1, include=FALSE}
names(data.Import)
# [1] "beta"      "M"         "intensity" "Meth"      "Unmeth"    "detP"     
# [7] "B.count"   "pd"        "TAG"       "minfi.Set"


dim(data.Import$beta)
# [1] 937055      3

data.Import$beta[1:4,1:4]
#                   Sample1    Sample2    Sample3
#cg25335775_BC21 0.09735627 0.09305994 0.07531023
#cg25341653_BC21 0.84000000 0.84889241 0.86613845
#cg25372195_BC21 0.11298640 0.28216165 0.20000000
#cg25411725_BC21 0.80538618 0.78527185 0.81116158

data.Import$minfi.Set$rgSet
#class: RGChannelSet 
#dim: 1105209 3 
#metadata(0):
#assays(2): Green Red
#rownames(1105209): 1600157 1600179 ... 99810982 99810990
#rowData names(0):
#colnames(3): Sample1 Sample2 Sample3
#colData names(6): Sample_Name Slide ... Sex Sample_Group
#Annotation
#  array: IlluminaHumanMethylationEPICv2
#  annotation: 20a1.hg38

data.Import$minfi.Set$mSet
#class: MethylSet 
#dim: 937055 3 
#metadata(0):
#assays(2): Meth Unmeth
#rownames(937055): cg25335775_BC21 cg25341653_BC21 ...
#  nv-GRCh38-chr7-152265113-152265113-C-A_BC11
#  nv-GRCh38-chr7-152265113-152265113-C-T_TC11
#rowData names(0):
#colnames(3): Sample1 Sample2 Sample3
#colData names(6): Sample_Name Slide ... Sex Sample_Group
#Annotation
#  array: IlluminaHumanMethylationEPICv2
#  annotation: 20a1.hg38
#Preprocessing
#  Method: NA
#  minfi version: NA
#  Manifest version: NA
```


## Generate `Address_chain_file`

`Address chain file` can be generated using `MeCall.SetChainFile`.
```{r Chunk2, include=FALSE}
MeCall.SetChainFile(remove.list = NULL, auto.remove = FALSE, global.env = TRUE)
```

If the remove.list argument is set to NULL, the chain file will be generated based on preset data implemented in MethylCallR.
`Duplicated.Probes.preset` is a data frame object that has a probe list to remove. This data is a subset of the Illumina EPICv2 manifest (`IlluminaHumanMethylationEPIC.V2a1.hg38`).

```{r Chunk3, include=FALSE}
dim(Duplicated.Probes.preset)
# [1] 6963   28

head(Duplicated.Probes.preset)[,1:4]
#                      Name AddressA_ID AddressB_ID Color_Channel
#cg00002033_TC12 cg00002033    22717226    17691929           Grn
#cg00002426_BC22 cg00002426    53715306          NA              
#cg00012692_BC12 cg00012692    60633427    68734181           Red
#cg00016699_BC22 cg00016699    35649404          NA              
#cg00017461_BC12 cg00017461    50639583    31712553           Red
#cg00021275_BC22 cg00021275    77698555          NA              
```

After running MeCall.SetChainFile, the Address_chain_file will be created as a global variable in R.

```{r Chunk4, include=FALSE}
ls()
# [1] "Address_chain_file" "data.Import"  

head(Address_chain_file)
#                EPICv2_Name EPICv2_AddressA_ID EPICv2_AddressB_ID
#cg25324105_BC11  cg25324105            1754126           99753217
#cg25383568_TC11  cg25383568           79792482           69667133
#cg25455143_BC11  cg25455143           80699190            7659147
#cg25459778_BC11  cg25459778           60797262           65710482
#cg25487775_BC11  cg25487775            5799427           89606481
#cg25595446_BC11  cg25595446           65640459           39619855
#                EPICv2_Color_Channel EPICv2_Infinium_Design_Type EPICv1_Name
#cg25324105_BC11                  Red                           I  cg25324105
#cg25383568_TC11                  Grn                           I  cg25383568
#cg25455143_BC11                  Red                           I  cg25455143
#cg25459778_BC11                  Grn                           I  cg25459778
#cg25487775_BC11                  Red                           I  cg25487775
#cg25595446_BC11                  Grn                           I  cg25595446
#                EPICv1_AddressA_ID EPICv1_AddressB_ID EPICv1_Color_Channel
#cg25324105_BC11           45678993           77772910                  Red
#cg25383568_TC11           80765370           80740275                  Grn
#cg25455143_BC11           32623209           16783987                  Red
#cg25459778_BC11           94603887           21788545                  Grn
#cg25487775_BC11           99729178           49657529                  Red
#cg25595446_BC11            1788109           90771247                  Grn
#                EPICv1_Infinium_Design_Type Methyl450K_Name
#cg25324105_BC11                           I      cg25324105
#cg25383568_TC11                           I      cg25383568
#cg25455143_BC11                           I      cg25455143
#cg25459778_BC11                           I      cg25459778
#cg25487775_BC11                           I      cg25487775
#cg25595446_BC11                           I      cg25595446
#                Methyl450K_AddressA_ID Methyl450K_AddressB_ID
#cg25324105_BC11               15610459               30645456
#cg25383568_TC11               18609425               33603340
#cg25455143_BC11               10691475               38761410
#cg25459778_BC11               56785457               15747493
#cg25487775_BC11               20602342               32758500
#cg25595446_BC11               27638328               41748394
#                Methyl450K_Color_Channel Methyl450K_Infinium_Design_Type
#cg25324105_BC11                      Red                               I
#cg25383568_TC11                      Grn                               I
#cg25455143_BC11                      Red                               I
#cg25459778_BC11                      Grn                               I
#cg25487775_BC11                      Red                               I
#cg25595446_BC11                      Grn                               I            
```

EPICv2 data will be converted based on this data.

On the other hand, users can provide a list of duplicated probes to be removed through the `RemoveDup` parameter in `MeCall.Filtering`.
If the user sets `"MethylCallR"` for the `RemoveDup` parameter, the preset data will be used.

```{r Chunk5, include=FALSE}
data.filtered <- MeCall.Filtering(data = data.Import, RemoveDup = "MethylCallR", badSample.minfi = FALSE)
```

To use user customized data, a data.frame object containing the probes to be removed as row names can be provided to the RemoveDup parameter. Ensure that the address ID information (named as `AddressA_ID` and `AddressB_ID`) for each probe is included in a column to create an `address chain file`. We recommend using a subset of the Illumina EPICv2 manifest (`data(IlluminaHumanMethylationEPIC.V2a1.hg38)`).

## Convert EPICv2 data to other versions

When converting a beta-value matrix from EPICv2 to EPICv1 :
```{r Chunk6, include=FALSE}
EPICv1.converted.Beta <- MeCallR.ShiftArray(data.filtered$beta, From = "EPICv2", To = "EPICv1", type = c("Matrix"))
```

```{r Chunk6-1, include=FALSE}
dim(EPICv1.converted.Beta)
# [1] 673840      3
```

When converting a RGChannelSet from EPICv2 to EPICv1 :
```{r Chunk7, include=FALSE}
EPICv1.converted.rgSet <- MeCallR.ShiftArray(data.filtered$minfi.Set$rgSet, From = "EPICv2", To = "EPICv1", type = c("rgSet"))
```

```{r Chunk7-1, include=FALSE}
EPICv1.converted.rgSet
#class: RGChannelSet 
#dim: 832127 3 
#metadata(0):
#assays(2): Green Red
#rownames(832127): 30652921 73600589 ... 73784382 73794434
#rowData names(0):
#colnames(3): Sample1 Sample2 Sample3
#colData names(6): Sample_Name Slide ... Sex Sample_Group
#Annotation
#  array: IlluminaHumanMethylationEPIC
#  annotation: ilm10b4.hg19
```

When converting a MethylSet from EPICv2 to EPICv1 :
```{r Chunk8, include=FALSE}
EPICv1.converted.mSet <- MeCallR.ShiftArray(data.filtered$minfi.Set$mSet, From = "EPICv2", To = "EPICv1", type = c("mSet"))
```

```{r Chunk8-1, include=FALSE}
EPICv1.converted.mSet
#class: MethylSet 
#dim: 722756 3 
#metadata(0):
#assays(2): Meth Unmeth
#rownames(722756): cg25324105 cg25383568 ... ch.12.78471492F
#  ch.21.43742285F
#rowData names(0):
#colnames(3): Sample1 Sample2 Sample3
#colData names(6): Sample_Name Slide ... Sex Sample_Group
#Annotation
#  array: IlluminaHumanMethylationEPIC
#  annotation: ilm10b4.hg19
#Preprocessing
#  Method: Raw (no normalization or bg correction)
#  minfi version: 1.48.0
#  Manifest version: 1.0.0
```






















